Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-12-25 thesis/article.tex
Thesis: Final fixes in Chapters 1 and 2, fixes from Th. Lianeas notes
Extended Abstract, Chapter 1 note fixes, Chapter 2 final checks and fixes, overall typos and rewordings

\thispagestyle{empty}
\pagenumbering{gobble}
\chapter*{Abstract}
In this masters thesis we study a $2$-level optimisation problem where, on an $n$-link network from a source $s$ to a target $t$, a unit of flow wants to move from $s$ to $t$ using the link with the lowest cost, while the link owners compete for profit by assigning tolls to links and thus creating a toll congestion game.
We examine only affine latencies for the links.
In addition, each flow player responds to tolls in a heterogeneous way, i.e. each flow player $p$ has a different money-time sensitivity value, which is described by a distribution function $\alpha(p)$.
First we present some basic definitions and properties of profit homogeneous games.
Using them as base, we examine cases of fixed and step distribution functions, where for the latter we first restrict our model to $2$-link instances and then we present strict conditions for the existence or not of a Nash Equilibrium.
We then introduce a new term, the split function $\as(t)$, defined as the time-money sensitivity value for which the $2$ link costs are equal for a given set of tolls $t$. With the help of $\as$ we will describe and prove some properties of the heterogeneous games for continuous distribution functions.
In our main result, we generalise the conditions necessary for the existence of a Nash Equilibrium in the profit game between two toll owners, and finally we discuss some extensions to $n$-link parallel networks.
\clearpage

@@

\thispagestyle{empty}
\pagenumbering{gobble}
\chapter*{Συνοψη}
Σε αυτή τη διπλωματική μεταπτυχιακού μελετάμε ένα πρόβλημα βελτιστοποίησης $2$ επιπέδων, όπου σε ένα δίκτυο με $n$ ακμές από μια πηγή $s$ σε ένα στόχο $t$, μια μονάδα ροής θέλει να μετακινηθεί από το $s$ στο $t$ χρησιμοποιώντας την ακμή με το χαμηλότερο κόστος, ενώ οι δύο ιδιοκτήτες των ακμών ανταγωνίζονται για κέρδος βάζοντας δίοδια στις ακμές και δημιουργώντας έτσι ένα παίγνιο συμφόρησης με διόδια.
Εξετάζουμε μόνο αφινικές καθυστερήσεις για τις ακμές.
Επιπλέον, ο κάθε παίκτης ροής αντιδρά στα διόδια με ετερογενή τρόπο, δηλ. ο κάθε παίκτης ροής $p$ έχει διαφορετική τιμή ευαισθησίας χρόνου-χρήματος, η οποία περιγράφεται από μια συνάρτηση κατανομής $\alpha(p)$.
Πρώτα παρουσιάζουμε κάποιους βασικούς ορισμούς και ιδιότητες των ομογενών παιγνίων κέρδους.
Mε αυτά ως βάση, εξετάζουμε περιπτώσεις σταθερών και βηματικών συναρτήσεων κατανομής, όπου για το τελευταίο πρώτα περιοριζόμαστε σε στιγμιότυπα $2$ ακμών και παρουσιάζουμε αυστηρές συνθήκες για την ύπαρξη ή μη σημείου ισορροπίας Nash.
Μετά εισαγάγουμε έναν νέο όρο, τη συνάρτηση διαχωρισμού $\as(t)$, ορισμένη ως η τιμή ευαισθησίας χρόνου-χρήματος για την οποία τα κόστη των $2$ ακμών είναι ίσα για ένα δοσμένο σετ διοδίων $t$.
Με τη βοήθεια της $\as$ θα περιγράψουμε και θα αποδείξουμε μερικές ιδιότητες των ετερογενών παιγνίων με συνεχή συνάρτηση κατανομής.
Στο κύριο αποτέλεσμά μας, γενικοποιούμε τις συνθήκες που είναι απαραίτητες για την ύπαρξη ισορροπίας Nash στο παίγνιο κέρδους μεταξύ δύο ιδιοκτητών ακμών, και τελικώς συζητάμε κάποιες επεκτάσεις για δίκτυα με $n$ ακμές.
\clearpage

@@

In this thesis we study a two level optimisation problem.
On the basis there is a network with 2 nodes $(s, t)$ and $n$ links, each link directed from $s$ to $t$ and weighted by a non-decreasing latency function that defines how the traffic increases as more players flow into that link.
On the first level, a flow $[0, 1]$ wants to move from $s$ to $t$ using one of the $n$ links and experience minimum latency (traffic).
On the second level, each link is owned by a different owner who can add a toll to their link and gain profit proportional to the flow that uses it.
The tolls impose an additional cost to the flow players, affecting the first level game, and the link owners compete with each other to maximise their profit.
In our case we will additionally assume that the flow users behave in a heterogeneous way when it comes to tolls, meaning that each player has a different money-time trade-off and thus see the link costs differently.

@@

Entering the main body of our work, in Chapter \ref{chapter:split} we introduce a new term regarding any two links of the network, called \textit{money-time sensitivity split}, which captures the notion of a (potentially non-existent) player whose money-time sensitivity makes them view the two links as equal.
We will then analyse that notion and provide properties and bounds for its values.
Finally, in the chapter's main result, we prove a general property of homogeneous network games as, due to the previously found bounds, there might be a portion of users whose sensitivity value can be arbitrarily large without affecting the game.

Having arrived at our main results in this work, in Chapter \ref{chapter:pricing_equilibria} we investigate the pricing competition game between the link owners of a network with heterogeneous users.
We begin by examining a class of games where the distribution function is fixed across all users, which, as we'll formally describe, makes the game \textit{pseudo-heterogeneous}.
Particular interest in the above class is the relation between games with identical setup but with different fixed distribution functions.
We describe and prove some properties for them, and using those we then tackle step distribution functions (where the user sensitivity can take values out of a finite set).
At this point we restrict our model to instances with $2$ links, and provide strict restrictions for equilibrium existence for the pricing competition game.

@@

Selfish routing has a vast history of literature as it's been proven useful in solving a variety of real-world problems in economics \cite{pigou1920economics}, transportation \cite{beckmann1956studies} and more (see \cite{roughgarden2002sr} and \cite{roughgarden2005slpoa} for more references).
It has been well-established, at least as early as 1920 from the work of Pigou \cite{pigou1920economics}, that selfish behavior in congested networks can create (arbitrary) inefficiencies with regard to the total latency experienced by all users.
A long-standing practice to solve these inefficiencies has been to regulate the network by adding tolls to the links, thus altering the cost of the link usage.
A classical result on this is the use of marginal tolls (Beckman et al. \cite{beckmann1956studies}), where each link user is charged a toll corresponding to their externality, i.e. the added congestion effect caused by their participation in the network.
This technique can eradicate the inefficiencies and create optimal flow in the network, however it assumes a strong homogeneity among the users.
Marginal cost pricing has been investigated in networks with homogeneous users (e.g. Dafermos \cite{dafermos1973toll}), with the resulting payment required by users on same links being different, an unwanted and non-practical approach, which also requires knowledge of the users' sensitivity beforehand.

@@

Regarding the model properties, there is a multitude of variations in the classical selfish routing model.
Beginning with the network topology, one can distinguish between parallel networks (our case) or series-parallel ones.
In addition the flow traversing through the network can either be atomic (finite number of players, each contributing to congestion equally or weighted) or non-atomic (infinite number of players, each contributing to congestion by an non-influencing infinitesimal amount).
Also the model might admit elastic demand (e.g. \cite{10.1287/moor.1060.0231}, \cite{Hearn1998}), a property where the users have a bound on the cost they're willing to pay, making instances having potential players not participating in the flow.

Most work considers homogeneous network users.
Heterogeneity has been studied by Schmeilder \cite{1973JSP.....7..295S} and Milchtaich \cite[Prop 3.3]{doi:10.1287/moor.25.3.349.12220} has contributed through with work on the more general crowding games, a class of games where the cost assigned to each player is affected only by the number of players selecting the same action (or strategy).
Large crowding games are basically less restrictive $n$-link parallel congestion games.
Also, Cole et al. \cite{10.1145/780542.780618} in their work showed that even general heterogeneous networks can be priced so that an optimal routing emerges while computing those efficient tolls can be done in polynomial time for convex latency functions and distribution functions with only finitely many values.
The latter paper has been one of the main pillars of this thesis.

@@

Acemoglu and Ozdaglar were the first to introduce pricing competition between link owners to the above model \cite{10.1287/moor.1060.0231}.
They showed that increasing the competition among operators from a monopoly to an oligopoly may reduce the efficiency of the network, achieving tight bounds on the Price of Anarchy.
In a follow-up work \cite{10.1109/JSAC.2007.070812}, they generalised the above by assuming more general topologies where links can also be linked serially (series-parallel networks).
Correa et al. \cite{correa2018pricing} showed pricing games may not exist, may not be unique and can be arbitrarily inefficient, but regulating the network with toll caps can solve all three issues.
Harkes et al. \cite{Harks_2019} then showed that the same is true for uniform toll caps, which is stronger in practice as toll discrimination is often not allowed.
The latter two papers have also been main pillars of the current work.

@@

In this section we will describe the model utilised throughout this thesis.
If we attempted to call that model by including all of its properties, the result would be a Heterogeneous Non-atomic 2-link Parallel Network Toll Congestion Pricing Game, so while we will describe each of those properties, we will eventually focus on heterogeneity and the pricing competition, thus calling it a Heterogeneous Pricing Game.

@@

A \textit{flow} is a Lebesque measurable function $f: [0, 1] \rightarrow N$ that describes which link is selected by each player.
It is more intuitive, however, to consider the resulting flow on each link (\textit{flow on paths}), i.e. $x_i = \lambda(\{p \in [0, 1]: f(p) = i\})$.
Hence we get flow as a (stochastic) vector $x = (x_i)_{i \in N}$ where $x_i$ is the total flow on link $i$ with $x_i \geq 0$ and $\sum_{i \in N}x_i = 1$.
The flow then creates congestion on the links, each described by a non-decreasing latency function $(\ell_i)_{i \in N}$ which we assume to be affine.
We denote by $\mathcal{L}_d$ the class of polynomial latency functions with nonnegative coefficients and degree at most $d$, and as such $\mathcal{L}_1$ becomes the class of affine latency functions.
Finally, the network can also be extended by allowing a set of tolls $t = (t_i)_{i \in N}$ to be assigned to each link, which in turn adds to the effective cost of a player using them.
At this point we should also mention that we will freely use, when needed, the standard game-theoretical notation of $t_{-i} = t \setminus \{t_i\}$.

@@

We can therefore denote by $x(t)$ the flow occurring on that unique equilibrium for a given toll $t$.
If $t = 0^N$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that looking at Lemma \ref{lemma:wardrop_equilibrium}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls, and not their actual values.
To demonstrate this, if we consider an arbitrary $c \in \R_+$ and tolls $t^\prime = t + c$ then for the same flow $x$ and any other flow $x^\prime$ the sum becomes
\begin{flalign*}
	\sum_{i \in N} (\ell_i(x_i) + t_i + c) (x_i - x_i^\prime) &= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) + c \left(\sum_{i \in N} x_i - \sum_{i \in N} x_i^\prime\right) \\
	&= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) \leq 0
\end{flalign*}
which makes $x$ the Wardrop equilibrium for $t^\prime$ as well.
The last equality holds because $x_i, x_i^\prime$ are flows of the network and therefore $\sum_{i \in N} x_i = \sum_{i \in N} x_i^\prime = 1$.
Also, this reasoning is still valid for $c < 0$ as long as $t_i + c \ge 0$ for all $i \in N$, or in other words, as long as $c \ge -\min_{i \in N}{t_i}$.

@@

Finally for affine latencies we can calculate $x(t)$.
For $\ell_i(x_i) = a_i x_i + b_i$ with $a_i > 0, b_i \geq 0$ and $t \in \R_+^N$, define $N(t) = \{i \in N | x_i(t) > 0\}$.
Since $x(t)$ is an equilibrium for $t$ then for all $i \in N(t)$ we get $\ell_i(x_i) + t_i = K$ for some common effective cost $K$.
By also using the fact that $\sum_{j \in N(t)} x_i = 1$ we solve the equations with regard to $K$ and eventually get for all $i \in N(t)$
\begin{equation}
	\label{eq:homogeneous_x_i}
	x_i(t) = \frac{1 + \sum_{j \in N(t)}\frac{b_j + t_j - b_i - t_i}{a_j}}{\sum_{j \in N(t)}\frac{a_i}{a_j}}
\end{equation}

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money-time sensitivity values.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(t) = \ell_i(x_i(t)) + \alpha(p) \cdot t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
%Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Lemma \ref{lemma:a_0_0}, allowing many players close to $0$ breaks the pricing game.

We can therefore define an instance of a Heterogeneous Pricing Game as the tuple $(N, \ell, \alpha)$ with $N = \{1, 2, \dots, n\}$, $\ell = (\ell_i)_{i \in N}$ and $\alpha$ a non-decreasing distribution function.
Then with all the above we can again define a Nash equilibrium for the game if $\forall i \in N$ and $\forall p \in [0, 1]$ it holds that
\[c_{f(p)}^p(t) \leq c_i^p(t)\]
Existence of an equilibrium is guaranteed by the more general results of Schmeidler \cite[Thm 2]{1973JSP.....7..295S}, while uniqueness is covered by Milchtaich \cite[Prop 3.3]{doi:10.1287/moor.25.3.349.12220} (see Related Work \ref{section:related_work}); for a more detailed coverage of those look at the definitions and propositions of Cole et al. \cite[\S2]{10.1145/780542.780618}.
We can therefore also extend the definition of $x(t)$, being the Nash Equilibrium for $t$, similarly as in the homogeneous case.

Finally, we will prove here a general property of parallel networks which capture the relationship between latencies and tolls.
It is trivial for homogeneous games that higher tolls create lower latencies on the respective edges, but for the heterogeneous case it is not straightforward so we'll prove it.

\begin{lemma}
	\label{lemma:latencies_tolls}
	For a Heterogeneous Parallel Game $(N, \ell, \alpha)$ with tolls $t$ and flow $x(t)$ the Nash equilibrium for $t$, it holds for all $i, j \in N$ with $x_i(t), x_j(t) > 0$ that
	\begin{enumerate}[(i)]
		\item $\ell_i(x_i(t)) < \ell_j(x_j(t))$ iff $t_i > t_j$
		\item $\ell_i(x_i(t)) = \ell_j(x_j(t))$ iff $t_i = t_j$
		\item $\ell_i(x_i(t)) > \ell_j(x_j(t))$ iff $t_i < t_j$
	\end{enumerate}
\end{lemma}

\begin{proof}
	We will prove $(i)$ and the rest follow similarly.
	For the left-to-right direction we assume that $\ell_i(x_i(t)) < \ell_j(x_j(t))$.
	If now $t_i \le t_j$ then for all players $p$ we get $\ell_i(x_i(t)) + \alpha(p) \cdot t_i < \ell_j(x_j(t)) + \alpha(p) \cdot t_j \Rightarrow c_i^p(t) < c_j^p(t)$.
	Since $x_j(t) > 0$ there exist players $p_j$ on link $j$ for which $c_{f(p_j)}^{p_j}(t) = c_j^{p_j}(t) > c_i^{p_j}(t)$, which is a contradiction since $x(t)$ is a Nash equilibrium.
	The opposite direction is also proved similarly.
\end{proof}

This property shows us that for a give $t \in \R_+^N$, if we order the links according to decreasing tolls, then the latencies are ordered in increasing order (and vice-versa).
It's easy to show that the players in each link are also ordered according to their $\alpha$ value increasing.
Therefore tolls define an ordering $t_1 \ge t_2 \ge \dots t_n$ where also $\ell_1(x_1(t)) \le \ell_2(x_2(t)) \le \dots \le \ell_n(x_n(t))$ and if we select player representatives from each link $p_1, p_2, \dots, p_n$ we also get $\alpha(p_1) \le \alpha(p_2) \le \dots \le \alpha(p_n)$.
We continue with the follow lemma which describes a property for the last link in an ordering.

>>>>

\thispagestyle{empty}
\pagenumbering{gobble}
\chapter*{Abstract}
In this masters thesis we study a $2$-level game where, on an $n$-link network from a source $s$ to a target $t$, a unit of flow wants to move from $s$ to $t$ using the link with the lowest cost, while the link operators compete for profit by assigning tolls to links and thus creating a toll congestion game.
We examine only affine latencies for the links.
In addition, each flow user responds to tolls in a heterogeneous way, i.e. each player $p$ in the flow has a different money-time sensitivity value, which is described by a distribution function $\alpha(p)$.
We first present some definitions and properties of pricing homogeneous games, heavily based on previous work, and using them as base, we extend them to describe pricing games with heterogeneous users.
We introduce a new term, the sensitivity split function $\as(t)$, defined for any $2$ links of the network as the time-money sensitivity value for which, given a set of tolls $t$, the link costs are equal.
With the help of $\as$ we describe some properties around the selfish behaviour of heterogeneous users for different tolls, as well as prove the existence of bounds for the sensitivity values that are relevant to the game.
Then, arriving at the main body of this study, we investigate the existence of a Nash Equilibrium in the profit game between the link operators for different cases of heterogeneous users.
We show there is no Nash Equilibrium when there are many users with $0$ sensitivity, i.e. players not affected by tolls, so we further assume $\alpha(p) > 0$.
We then investigate cases of fixed distribution functions, which, even though create homogeneous behaviour for the users (making them pseudo-heterogeneous games), help us describe how flow, tolls and equilibria behave as that fixed value changes.
Using all the above, we study step distribution functions, where by limiting our scope to $2$ links, we present strict conditions necessary for the existence or not of a Nash Equilibrium in the profit game between the link operators.
We present and analyse an algorithm to calculate that equilibrium, if it exists, and finish with some discussion about potential future work.
\clearpage

@@

\thispagestyle{empty}
\pagenumbering{gobble}
\chapter*{Συνοψη}
Σε αυτή τη διπλωματική μεταπτυχιακού μελετάμε ένα παίγνιο $2$ επιπέδων όπου, σε ένα δίκτυο με $n$ ακμές από μια πηγή $s$ σε ένα στόχο $t$, μια μονάδα ροής θέλει να μετακινηθεί από το $s$ στο $t$ χρησιμοποιώντας την ακμή με το χαμηλότερο κόστος, ενώ οι διαχειριστές των ακμών ανταγωνίζονται για κέρδος βάζοντας δίοδια στις ακμές και δημιουργώντας έτσι ένα παίγνιο συμφόρησης με διόδια.
Εξετάζουμε μόνο αφινικές καθυστερήσεις για τις ακμές.
Επιπλέον, ο κάθε χρήστης δικτύου αντιδρά στα διόδια με ετερογενή τρόπο, δηλ. ο κάθε παίκτης $p$ στη ροή του δικτύου έχει διαφορετική τιμή ευαισθησίας χρόνου-χρήματος, η οποία περιγράφεται από μια συνάρτηση κατανομής $\alpha(p)$.
Πρώτα παρουσιάζουμε κάποιους βασικούς ορισμούς και ιδιότητες των ομογενών παιγνίων κέρδους, κυρίως βασισμένα σε προηγούμενες δουλειές, και με αυτά ως βάση, τα επεκτείνουμε για να περιγράψουμε παίγνια κέρδους με ετερογενείς χρήστες.
Εισαγάγουμε έναν νέο όρο, τη συνάρτηση διαχωρισμού ευαισθησίας $\as(t)$, ορισμένη για οποιεσδήποτε $2$ ακμές του δικτύου ως η τιμή ευαισθησίας χρόνου-χρήματος για την οποία, δωσμένων σετ διοδίων $t$, τα κόστη των $2$ ακμών είναι ίσα.
Με τη βοήθεια της $\as$ θα περιγράψουμε μερικές ιδιότητες γύρω από την εγωιστική συμπεριφορά των ετερογενών χρηστών για διαφορετικά διόδια, καθώς και θα αποδείξουμε την ύπαρξη άνω φράγματος για τις τιμές ευαισθησίας που αφορούν το παίγνιο.
Στη συνέχεια, φτάνοντας στο κύριο σώμα αυτής της έρευνας, εξετάζουμε την ύπαρξη ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών για διαφορετικές περιπτώσεις ετερογενών χρηστών.
Δείχνουμε ότι δεν υπάρχει ισορροπία Nash όταν υπάρχουν πολλοί χρήστες με $0$ ευαισθησία, δηλ. παίκτες που δεν επηρρεάζονται από τα διόδια, οπότε στο εξής υποθέτουμε ότι $\alpha(p) > 0$.
Κατόπιν εξετάζουμε περιπτώσεις όπου η συνάρτηση κατανομής είναι σταθερή, η οποίες, αν και δημιουργούν ομογενή συμπεριφορά στους χρήστες (κάνοντάς τα ψευδο-ετερογενή παίγνια), μας βοηθάνε να περιγράψουμε πώς συμπεριφέρονται η ροή, τα διόδια και τα σημεία ισορροπίας όταν αυτή η σταθερή τιμή μεταβάλλεται.
Χρησιμοποιώντας όλα τα παραπάνω, μελετάμε βηματικές συναρτήσεις κατανομής, όπου περιορίζοντας το μοντέλο μας σε στιγμιότυπα $2$ ακμών, παρουσιάζουμε αυστηρές συνθήκες αναγκαίες για την ύπαρξη ή μη σημείου ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών.
Παρουσιάζουμε και αναλύουμε έναν αλγόριθμο για τον υπολογισμό του σημείου ισορροπίας, αν υπάρχει, και κλείνουμε με μια συζήτηση για πιθανή μελοντική δουλειά.

\clearpage

@@

In this thesis we study a two level game.
On the basis there is a network with 2 nodes $(s, t)$ and $n$ links, each link directed from $s$ to $t$ and weighted by a non-decreasing latency function that defines how the traffic increases as more players flow into that link.
On the lower level, a flow $[0, 1]$ wants to move from $s$ to $t$ using one of the $n$ links and experience minimum latency (traffic).
On the upper level, each link is managed by a different operator who can assign a toll to their link and gain profit proportional to that toll and the flow that ends up using the link.
The tolls impose an additional cost to the flow players, affecting the lower level game, and the link operators compete with each other to maximise their profit.
In our case we will additionally assume that the flow users behave in a heterogeneous way when it comes to tolls, meaning that each player has a different money-time trade-off and thus see the link costs differently.

@@

Entering the main body of our work, in Chapter \ref{chapter:split} we introduce a new term regarding any two links of the network, called \textit{money-time sensitivity split}, which captures the notion of a (potentially non-existent) player whose money-time sensitivity makes them view the two links as equal.
We will then analyse that notion and provide properties and bounds for its values.
Finally, in the chapter's main result, we prove a general property of heterogeneous network games as, due to the previously found bounds, there might be a portion of users whose sensitivity value can be arbitrarily large without affecting the game.

Having arrived at our main results in this work, in Chapter \ref{chapter:pricing_equilibria} we investigate the pricing competition game between the link operators of a network with heterogeneous users.
We begin by examining a class of games where the distribution function is fixed across all users, which, as we'll formally describe, makes the game \textit{pseudo-heterogeneous}.
Particular interest in the above class is the relation between games with identical setup but with different fixed distribution functions.
We describe and prove some properties for them, and using those we then tackle step distribution functions (where the user sensitivity can take values out of a finite set).
At this point we restrict our model to instances with $2$ links, and provide strict restrictions for equilibrium existence for the pricing competition game.

@@

Selfish routing has a vast history of literature as it's been proven useful in solving a variety of real-world problems in economics \cite{pigou1920economics}, transportation \cite{beckmann1956studies} and more (see \cite{roughgarden2002sr} and \cite{roughgarden2005slpoa} for more references).
It has been well-established, at least as early as 1920 from the work of Pigou \cite{pigou1920economics}, that selfish behavior in congested networks can create (arbitrary) inefficiencies with regard to the total latency experienced by all users.
A long-standing practice to solve these inefficiencies has been to regulate the network by adding tolls to the links, thus altering the cost of the link usage.
A classical result on this is the use of marginal tolls (Beckman et al. \cite{beckmann1956studies}), where each link user is charged a toll corresponding to their externality, i.e. the added congestion effect caused by their participation in the network.
This technique can eradicate the inefficiencies and create optimal flow in the network, however it assumes a strong homogeneity among the users.
Marginal cost pricing has been investigated in networks with heterogeneous users (e.g. Dafermos \cite{dafermos1973toll}), with the resulting payment required by users on same links being different, an unwanted and non-practical approach, which also requires knowledge of the users' sensitivity beforehand.

@@

Regarding the model properties, there is a multitude of variations in the classical selfish routing model.
Beginning with the network topology, one can distinguish between parallel networks (our case) or series-parallel ones.
In addition the flow traversing through the network can either be atomic (finite number of players, each contributing to congestion equally or weighted) or non-atomic (infinite number of players, each contributing to congestion by an non-influencing infinitesimal amount).
Also the model might admit elastic demand (e.g. \cite{10.1287/moor.1060.0231}, \cite{Hearn1998}), a property where users have a bound on the cost they're willing to pay, making a player potentially not participate in the flow if all link costs are above that bound.

Most work considers homogeneous network users.
Heterogeneity has been studied by Schmeilder \cite{1973JSP.....7..295S} while Milchtaich \cite[Prop $3.3$]{doi:10.1287/moor.25.3.349.12220} has contributed through with work on the more general crowding games, a class of games where the cost assigned to each player is affected only by the number of players selecting the same action (or strategy).
Large crowding games are basically less restrictive $n$-link parallel congestion games.
Also, Cole et al. \cite{10.1145/780542.780618} in their work showed that even general heterogeneous networks can be priced so that an optimal routing emerges while computing those efficient tolls can be done in polynomial time for convex latency functions and distribution functions with only finitely many values.
The latter paper has been one of the main pillars of this thesis.

@@

Acemoglu and Ozdaglar were the first to introduce pricing competition between link operators to the above model \cite{10.1287/moor.1060.0231}.
They showed that increasing the competition among operators from a monopoly to an oligopoly may reduce the efficiency of the network, achieving tight bounds on the Price of Anarchy.
In a follow-up work \cite{10.1109/JSAC.2007.070812}, they generalised the above by assuming more general topologies where links can also be linked serially (series-parallel networks).
Correa et al. \cite{correa2018pricing} showed pricing games may not exist, may not be unique and can be arbitrarily inefficient, but regulating the network with toll caps can solve all three issues.
Harkes et al. \cite{Harks_2019} then showed that the same is true for uniform toll caps, which is stronger in practice as toll discrimination is often not allowed.
The latter two papers have also been main pillars of the current work.

@@

In this section we will describe the model utilised throughout this thesis.
If we attempted to call that model by including all of its properties, the result would be a Heterogeneous Non-atomic $n$-link Parallel Network Toll Congestion Pricing Game, so while we will describe each of those properties, we will eventually focus on heterogeneity and the pricing competition, thus calling it a Heterogeneous Pricing Game.

@@

A \textit{flow} is a Lebesque measurable function $f: [0, 1] \rightarrow N$ that describes which link is selected by each player.
It is more intuitive, however, to consider the resulting flow on each link (\textit{flow on paths}), i.e. $x_i = \lambda(\{p \in [0, 1]: f(p) = i\})$.
Hence we get flow as a (stochastic) vector $x = (x_i)_{i \in N}$ where $x_i$ is the total flow on link $i$ with $x_i \geq 0$ and $\sum_{i \in N}x_i = 1$.
The flow then creates congestion on the links, each described by a non-decreasing latency function $(\ell_i)_{i \in N}$ which we assume to be affine.
We denote by $\mathcal{L}_d$ the class of polynomial latency functions with nonnegative coefficients and degree at most $d$, and as such $\mathcal{L}_1$ becomes the class of non-decreasing latency functions.
Finally, the network can also be extended by allowing a set of tolls $t = (t_i)_{i \in N}$ to be assigned to each link, which in turn adds to the effective cost of a player using them.
At this point we should also mention that we will freely use, when needed, the standard game-theoretical notation of $t_{-i} = t \setminus \{t_i\}$.

@@

The notion of uniqueness under costs is important in order to acknowledge that a given flow on paths can be achieved by many player-link assignments, since the users can be put arbitrarily on the links, as long as each link is receiving the same total traffic.

We can therefore denote by $x(t)$ the flow occurring on that unique up to costs equilibrium for a given set of toll $t$.
If $t = 0^N$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that looking at Lemma \ref{lemma:wardrop_equilibrium}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls, and not their actual values.
To demonstrate this, if we consider an arbitrary $c \in \R_+$ and tolls $t^\prime = t + c$ then for the same flow $x$ and any other flow $x^\prime$ the sum becomes
\begin{flalign*}
	\sum_{i \in N} (\ell_i(x_i) + t_i + c) (x_i - x_i^\prime) &= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) + c \left(\sum_{i \in N} x_i - \sum_{i \in N} x_i^\prime\right) \\
	&= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) \leq 0
\end{flalign*}
which makes $x$ the Wardrop equilibrium for $t^\prime$ as well.
The last equality holds because $x_i, x_i^\prime$ are flows of the network and therefore $\sum_{i \in N} x_i = \sum_{i \in N} x_i^\prime = 1$.
Also, this reasoning is still valid for $c < 0$ as long as $t_i + c \ge 0$ for all $i \in N$, or in other words, as long as $c \ge -\min_{i \in N}{t_i}$.

@@

Finally for affine latencies we can calculate $x(t)$ by following the same reasoning as \cite[Prop $3.1$]{Harks_2019}.
For $\ell_i(x_i) = a_i x_i + b_i$ with $a_i > 0, b_i \geq 0$ and $t \in \R_+^N$, define $N(t) = \{i \in N | x_i(t) > 0\}$.
Since $x(t)$ is an equilibrium for $t$ then for all $i \in N(t)$ we get $\ell_i(x_i) + t_i = K$ for some common effective cost $K$.
By also using the fact that $\sum_{j \in N(t)} x_i = 1$ we solve the equations with regard to $K$ and eventually get for all $i \in N(t)$
\begin{equation}
	\label{eq:homogeneous_x_i}
	x_i(t) = \frac{1 + \sum_{j \in N(t)}\frac{b_j + t_j - b_i - t_i}{a_j}}{\sum_{j \in N(t)}\frac{a_i}{a_j}}
\end{equation}

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money-time sensitivity values.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of an edge with flow $x_i$ and toll $t_i$ as $c_i(p) = \ell_i(x_i) + \alpha(p) \cdot t_i$.
For a flow $x$ and tolls $t$, each player considers their costs experienced across all links and then selects to use the one (or one of potentially many) with the lowest cost. 
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, +\infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
%Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Lemma \ref{lemma:a_0_0}, allowing many players close to $0$ breaks the pricing game.

We can therefore define an instance of a Heterogeneous Pricing Game as the tuple $(N, \ell, \alpha)$, with $N = \{1, 2, \dots, n\}$ the links of the network, $\ell = (\ell_i)_{i \in N}$ the affine latency functions for each link and $\alpha$ the non-decreasing distribution function.
With all the above we can again define a Nash equilibrium for the game if $\forall i \in N$ and $\forall p \in [0, 1]$ it holds that
\[c_{f(p)}(p) \leq c_i(p)\]
Existence of an equilibrium is guaranteed by the more general results of Schmeidler \cite[Thm 2]{1973JSP.....7..295S}, while uniqueness with regard to costs is covered by Milchtaich \cite[Prop 3.3]{doi:10.1287/moor.25.3.349.12220} (see Related Work \ref{section:related_work}); for a more detailed coverage of those look at the definitions and propositions of Cole et al. \cite[\S2]{10.1145/780542.780618}.
We can therefore also extend the definition of $x(t)$, being the Nash Equilibrium for $t$, similarly as in the homogeneous case.

We will now prove a general property of parallel networks which captures the relationship between latencies and tolls.
It is trivial for homogeneous games that higher tolls create lower latencies on their respective links, but for the heterogeneous case it is not straightforward so we'll prove it.

\begin{lemma}
	\label{lemma:latencies_tolls}
	Let $(N, \ell, \alpha)$ a Heterogeneous Parallel Game with tolls $t$ and flow $x(t)$ the Nash equilibrium for $t$.
	It holds for all $i, j \in N$ with $x_i(t), x_j(t) > 0$ that
	\begin{enumerate}[(i)]
		\item $\ell_i(x_i(t)) < \ell_j(x_j(t))$ iff $t_i > t_j$
		\item $\ell_i(x_i(t)) = \ell_j(x_j(t))$ iff $t_i = t_j$
		\item $\ell_i(x_i(t)) > \ell_j(x_j(t))$ iff $t_i < t_j$
	\end{enumerate}
\end{lemma}

\begin{proof}
	We will prove $(i)$ and the rest follow similarly.
	For the left-to-right direction we assume that $\ell_i(x_i(t)) < \ell_j(x_j(t))$.
	If we also assume that $t_i \le t_j$, then for all players $p$ we get $\ell_i(x_i(t)) + \alpha(p) \cdot t_i < \ell_j(x_j(t)) + \alpha(p) \cdot t_j \Rightarrow c_i(p) < c_j(p)$.
	Since $x_j(t) > 0$ there exist players $p_j$ on link $j$ for which $c_{f(p_j)}(p_j) = c_j(p_j) > c_i(p_j)$, which is a contradiction since $x(t)$ is a Nash equilibrium.
	The opposite direction is also proved similarly.
\end{proof}

This property shows us that for a given $t \in \R_+^N$, if we order the links according to decreasing toll values, then the same link order will have latencies ordered in increasing values (and vice-versa).
It is then easy to show that the players using each such link are also ordered with regard to their $\alpha$ values in increasing order.
Therefore a given set of tolls can define an ordering of links where $t_1 \ge t_2 \ge \dots \ge t_n$ and $\ell_1(x_1(t)) \le \ell_2(x_2(t)) \le \dots \le \ell_n(x_n(t))$, while if we were to pick player representatives from each link with positive flow $p_1, p_2, \dots, p_k$, we also get $\alpha(p_1) \le \alpha(p_2) \le \dots \le \alpha(p_k)$.
Note that the ordering is unique only when the inequalities are strict, as had we had $t_i = t_j$ for any $i, j \in N$ then $i$ and $j$ can be ordered arbitrarily.
If $t = 0$ then all links can be ordered arbitrarily, which is why when comparing $t$ with $0$ we will assume the ordering is the same for simplicity.
We continue with the following lemma which describes a property for the last link in such an ordering.
